---
title: "Homework 2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

Proceed as in Homework 1. Solve the exercises and upload your code and explanations to SUCOURSE.

1.  Load the `dengue.csv` file provided to you. A description of the variables is also available.

    ```{r}
    library(here)
    library(readr)
    library(modelsummary)
    library(dplyr) 
    ```

    ```{r}
    dengue <- read_csv("dengue.csv")
    ```

Store the data set as `dengue`

------------------------------------------------------------------------

2.  Estimate a linear regression using average humidity to predict whether dengue was observed in the area, and look at the results.

-   Save the model as `m2`. Use the `msummary()` function from the **modelsummary** package to display the results with the `stars = TRUE` option.

    ```{r}
    m2 <- lm(NoYes~humid, # This is the linear model
              
              data = dengue)
    ```

    ```{r}
    msummary(m2, stars = TRUE)
    ```

------------------------------------------------------------------------

3.  Write a couple of sentences, one interpreting the intercept and one interpreting the slope.

    ```{r}
    # Intercept parameter: 
    # According to runned linear regression, intercept is -0.416 which means the regression line intercepts with Y axis at -0.416
    # intercept parameter is significant at more than 99% confidence level since p value is lower than 0.01 (even lower than 0.001)
    # Slope parameter:
    # According to runned linear regression, If humidity changes by 1 unit (%), there will be dengue with 0.05 more possibility.
    # Slope parameters is significant at more than 99% confidence level since p value is lower than 0.01 (even lower than 0.001)
    ```

    ------------------------------------------------------------------------

4.  Get a set of summary statistics for the humidity variable and write a comment on how this can help you make sense of the intercept in the regression from step 2.

    ```{r}
    summary(dengue$humid)
    # Intercept makes sense now since we have min 0.6714 humidity, which means humidity is always bigger than 0. To make the equation near 0, the intercept should be lower than 0.
    ```

You can use `summary()` on the variable.

------------------------------------------------------------------------

5.  We might recognize that, if we're interested in the [**effect of humidity on Dengue**]{.underline}, **temperature** might be on a back door. Add a control for temperature, rerun the regression, and show the results.

    ```{r}
    m5 = lm(NoYes~humid + temp, # This is the linear model
              
              data = dengue)
    ```

    ```{r}
    modelsummary(m5, stars = TRUE)
    # As one can see here, humid is still significantly important since it has p value 0.002 and same with intercept. Temperature affects dengue chance negatively when considered with humidity. Also, temperature is significantly important at 99% confidence level.
    # Even though temperature does not effect the equation much, we have eliminated endogeneity possibilities.
    ```

Save the model as `m5`. Use the `msummary()` function from the **modelsummary** package to display the results with the `stars = TRUE` option.

------------------------------------------------------------------------

6.  Our dependent variable is binary, and we're getting predictions below zero, which we might not want. **Rerun the regression from question 5 but as a logit model**, and report the [marginal effects of both slope coefficients]{.underline}.

    ```{r}
    m6 <- glm(NoYes~humid + temp, family=binomial(link="logit"), data=dengue) 
    ```

[**I know that we have not covered what the marginal effects are. Therefore, you will need to investigate this topic a bit to solve this question.**]{style="color: purple"}

Save the model as `m6`. In R, use the `msummary()` function from the **modelsummary** package to display the results with the `stars = TRUE` option.\

```{r}
modelsummary(m6, stars = TRUE)
# Marginal Effects of humid:
# If humid increases by 1 %, the probability of there is dengue increases by 0.3%
# Marginal Effects of temp:
# If tem increases by 1 degree, the probability of there is dengue increases by 0.04%
```

------------------------------------------------------------------------

7.  Now let's say we're directly interested in the relationship between temperature and humidity. Run an OLS regression of humidity on temperature. Calculate the residuals of that regression, and then make a plot that will let you evaluate whether there is likely heteroskedasticity in the model. Rerun the model with heteroskedasticity-robust standard errors. Show both models, and say whether you think there is heteroskedasticity

-   Save the model as `m7`, and `m7b` if you rerun with robust standard errors.

-   Use `fiter(!is.na(dengue$humid))` on the data before running the model so the residuals line up properly. Use the `msummary()` function from the **modelsummary** package to display the results with the `stars = TRUE` option. Also, if you're clever about `msummary()` usage, you can skip creating `m7b`. You can access residuals using `resid(m7)`.

    ```{r}
    filtered_dengue = dengue %>% filter(!is.na(dengue$humid))

    m7 <- lm(humid ~ temp, data = filtered_dengue)

    residuals <- resid(m7)

    plot(fitted(m7), residuals, ylab = "Residuals", xlab = "Fitted Values",
         main = "Residuals vs Fitted", pch = 16)

    m7b_robust <- lm(humid ~ temp, data = filtered_dengue, weights = 1 / residuals^2)

    msummary(list(m7, m7b_robust), stars = TRUE)
    # There should be heteroskedasticity since standard errors differs a lot and R^2 differs a lot. Also the plot shows that standard error is not around 0, and it differs a lot in specific areas. Since also the plot does not look like as expected, there is heteroskedasticity.
    ```

------------------------------------------------------------------------

8.  In the graph in the last problem you may have noticed that for certain ranges of temperate, the errors were clearly nonzero on average. This can indicate a functional form problem. Run the model from question 7 again (with heteroskedasticity-robust standard errors), but this time use the logarithm of humidity in place of humidity. Add a sentence interpreting the coefficient on temperature.

    ```{r}

    m8 <- lm(log(humid) ~ temp, data = filtered_dengue, weights = 1 / residuals^2)

    msummary(list(m8), stars = TRUE)
    # coefficient of temperature interpretation:
    # If temperature increases by 1 degrees, humidity increases by 0.05 unit (%).
    # It is significant at 99% confidence level since p value is nearly zero.
    ```

Store the model as `m8`. Use the `msummary()` function from the **modelsummary** package to display the results with the `stars = TRUE` option.
